---
title: "PG14_Backup"
author: "Farshad Ebrahimi"
date: '2022-11-22'
output: html_document
params:
  database: "mars14_data" 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Dplyr stuff
  library(magrittr)
  library(tidyverse)
  library(lubridate)

#Database Stuff
  library(odbc)

#Other stuff
  library(openssl)
  options(stringsAsFactors=FALSE)
  


```

```{r Section 1 - Back up the database to GSI Monitoring\07 Databases and Tracking Spreadsheets\18 MARS Database Back Up Files\PG 14, include = FALSE}
## Back up the database 2.0 ----  
#which database to backup? what format?
  format_archive <- "c"
  format <- paste("--format=",format_archive, sep = "")
  database_archive <-"mars_data"
  db <- paste("--dbname=",shQuote(database_archive), sep ="")
  
#specify other details: the pathway to find pg_dump, where to save archive, naming format, database server specs, and username credentials  
  datestring <- Sys.time() %>% format("%Y%m%dT%H%M")
  extension <- "pgdump"
  filename <- paste0(datestring, "_", "mars_data", ".", extension)
  filepath <- paste0("\\\\pwdoows\\oows\\Watershed Sciences\\GSI Monitoring\\07 Databases and Tracking Spreadsheets\\18 MARS Database Back Up Files\\PG 14\\", filename)
  filepath <- shQuote(filepath)
  pg_dump <- "C:\\Program Files\\pgAdmin 4\\v6\\runtime\\pg_dump.exe"
  pg_dump <- shQuote(pg_dump)
  host <- "PWDOOWSDBS"
  host <- shQuote(host)
  port <- "5434" 
  port <- shQuote(port)
  username <- "mars_admin"
  username <- shQuote(username)
  role <- "mars_admin"
  role <- shQuote(role)
  
  
#Assemble the entire pg_dump string	
  pgdumpstring <- paste(pg_dump,
                        "--file",filepath,
                        "--host",host,
                        "--port", port,
                        "--username",username,
                        "--no-password",
                        "--role",role,
                        format,
                        db)
 # run the command line using system function 
  results_dump <- system(pgdumpstring, intern = TRUE, wait = FALSE)

```


The `r params$database` database was `r ifelse(length(results_dump) == 0, "SUCCESSFULLY", "NOT")` backed up to the W drive.

`r if(length(results_dump) == 0){ paste("The file is located at", gsub("\\\\", "/", filepath)) }`

```{r Section 2 - Create Test DB, include = FALSE}	

## Create a database to host the test DB 3.0 ----  

  #create db
  #testdbname <- paste(database_archive,"_","archivetest_",datestring, sep = "") longnames don't get recognize by pg_restore
  con <- odbc::dbConnect(odbc::odbc(), params$database)

  error_db <- NULL
  tryCatch({
    
  testdatabase <- paste("db",Sys.time() %>% format("%Y%m%d"), sep = "_")
  query_str <- "CREATE DATABASE %s WITH TEMPLATE = template0 OWNER = mars_admingf"
  sql_query <- paste(sprintf(query_str,testdatabase),collapse="")
  results_createdb <- dbSendQuery(con, sql_query)
  
  },error=function(e){
  error_db <<- toString(conditionMessage(e))
  }
  )
  
  dbDisconnect(con)


```

```{r Section 3 - Restoring the  backup, include = FALSE}	

## Restore the archive ---- 4.0
  
  pg_restore <- "C:\\Program Files\\pgAdmin 4\\v6\\runtime\\pg_restore.exe"
  pg_restore <- shQuote(pg_restore)
  
  #Assemble the entire pg_restore string	
  pgrestorestring <- paste(pg_restore,
                        "--host",host,
                        "--port", port,
                        "--username",username,
                        "--no-password",
                        "--role",role,
                        "--jobs=5",
                        paste("--dbname=",shQuote(testdatabase), sep=""),
                        filepath)
  results_restore <- system(pgrestorestring, intern = TRUE, wait = FALSE)
  
```

The scratch database, was `r ifelse(length(results_createdb) == 0, "SUCCESSFULLY", "NOT")` created on the MARS postgres server.  
  
`r if(length(results_createdb) == 0){ paste("The database is named", testdatabase) }`  
  
The scratch database was `r ifelse(length(results_restore) == 0, "SUCCESSFULLY", "NOT")` populated by the database archive located at `r gsub("\\\\", "/", filepath)`

The scratch database will be deleted automatically if restoration has been successful. 

```{r Section 4 - Delete the scratch DB, include = FALSE, eval= length(results_restore)==0 & is.null(error_db)}	

# DB
  con <- odbc::dbConnect(odbc::odbc(), params$database)

  
  error_db_drop <- NULL
  tryCatch({
    
  query_str <- "DROP DATABASE %s"
  sql_query <- paste(sprintf(query_str,testdatabase),collapse="")
  db_dropped <- dbSendQuery(con, sql_query)
  
  },error=function(e){
  error_db_drop <<- toString(conditionMessage(e))
  }
  )
  
  dbDisconnect(con)
  
```

```{r Section 5 - Prune old backups, include = FALSE,eval= length(results_restore)==0 & is.null(error_db)}	
 
 ## Pruning old back ups
  #get a list of backup files  from the backup directory
  backups <- list.files("\\\\pwdoows\\oows\\Watershed Sciences\\GSI Monitoring\\07 Databases and Tracking Spreadsheets\\18 MARS Database Back Up Files\\PG 14")
  backup_pathway <- paste("\\\\pwdoows\\oows\\Watershed Sciences\\GSI Monitoring\\07 Databases and Tracking Spreadsheets\\18 MARS Database Back Up Files\\PG 14", backups, sep = "\\")
  
  #extract the backup date from the backup name and reformat it as Date
  backup_datestrings <- str_trunc(backups,8, "right", ellipsis = "")
  Dates <- as.Date(backup_datestrings, format="%Y%m%d")
  
  #add the weekdays
  backup_dates <-as.data.frame(Dates)
  backup_dates$W_Days <- weekdays(backup_dates$Dates)
  backup_dates$M_Days<- day(backup_dates$Dates)
  backup_dates$Days_Ago <- as.Date(Sys.time() %>% format("%Y-%m-%d"))-backup_dates$Dates
  
  #get index of those rows that are older than 7 days, and are not day 28 or Friday
  delete_lastmonth <- which(backup_dates$Days_Ago > 6 & backup_dates$Days_Ago < 30 & backup_dates$W_Days !="Friday" & backup_dates$M_Days !=28)
  delete_older <- which(backup_dates$Days_Ago > 29 & backup_dates$M_Days !=28)
  
  #prune the delete index 
  results_prune <- file.remove(backup_pathway[c(delete_lastmonth,delete_older)])
  
```

```{r Section 6 - Log Table, include = FALSE, eval= length(results_restore)==0}	

# DB
  con <- odbc::dbConnect(odbc::odbc(), params$database)
  time_stamp <- Sys.time()
  record_log <- data.frame(operation = c("backup","restore","prune"),
                           time = c(time_stamp,time_stamp,time_stamp),
                           backup_path = c(filepath,filepath,filepath),
                           success = c(ifelse(length(results_dump) == 0,1,0),
                                       ifelse(length(results_restore) == 0 & is.null(error_db),1,0),
                                       ifelse(length(results_prune) != 0, 1, 0)),
                           error = c(ifelse(length(results_dump) == 0, NA, results_dump),
                                       ifelse(length(results_restore) == 0 & is.null(error_db),NA,paste(results_restore, error_db)),
                                       ifelse(length(results_prune) != 0, NA,"No back up file was deleted!"))
                           )
  
  
  dbDisconnect(con)
  
```
